---
output:
  word_document: default
  html_document: default
---


Jeffrey Ryan Morris  
2/22/2021  
BAN 502  
Module 6: Assignment 1 Clustering



```{r include=FALSE}
library(tidyverse)
library(tidymodels)
```

```{r}
trucks = read.csv("trucks.csv")
```


### Task 1

It is unclear that any linear relationship exists between these variables. 
There appears to be 4 clusters. 
1) Distance 25 to 75, Speeding 17 to 50.
2) Distance 25 to 75, Speeding 0 to 17.
3) Distance 150 to 200, Speeding 0 to 25.
4) Distance ~110 and speeding ~ 75

```{r}
trucks %>% 
  ggplot(aes(x=Distance, y=Speeding)) +
  geom_point(alpha=0.4) + theme_bw()
```

### Task 2

```{r}
trucks = select(trucks, -Driver_ID)

kmeans_recipe = recipe(~ Distance + Speeding, trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks) #prepares the recipe

trucks_cleaned = bake(trucks_dummy, trucks) #applies the recipe and yields a data frame
```

### Task 3

K Mean cluster of 2 is not impressive for this problem.  There are obviously more clusters than 2 and I don't think a observation of 200 distance and 100 speeding is remotely similar to 250 distance and 10 speeding.

```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 2)
trucks_augmented1 = augment(clusters, trucks)
str(trucks_augmented1)
```
```{r}

ggplot(trucks_augmented1, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point()

```

### Task 4

A K value of 4 appears to be the best for this dataset.



```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```
```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

### Task 5

K of 4 appears to be the best.

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()
```

### Task 6
I think this K=4 result is nice. From the graph there appears to only be 1 data point that could be in the wrong group (Distance ~120, Speeding~5). Put otherwise pretty spot on
```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 4)
trucks_augmented2 = augment(clusters, trucks)
str(trucks_augmented2)
```

```{r}
ggplot(trucks_augmented2, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point()

```

