---
output:
  word_document: default
  html_document: default
---
Jeffrey Ryan Morris  
2/10/2021  
BAN 502  
Module 4: Assignment 2 Random Forests

```{r, include = FALSE}
library(tidyverse)
library(tidymodels)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(ranger) #for random forests
library(randomForest) #also for random forests
library(caret)
library(skimr)
library(GGally)
library(gridExtra)
library(vip) #variable importance
```

```{r}
drug = read.csv("drug_data-1.csv")
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)
```
```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"


drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>% mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)

```

```{r include=FALSE}
str(drug_clean)
```
```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)

```

### Task 1

There is no missingness for drug_clean

```{r}
skim(drug_clean)
```
### Task 2

```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) #70% in training
train = training(drug_split)
test = testing(drug_split)
```


### Task 3

Age has a strong impact on Nicotine use as younger people are more likely to use nicotine.
Gender has an impact on Nicotine use as more Men use nicotine than women.
Education has an impact on Nicotine it appears (though more analysis would be required) that less nicotine users are higher/advance in education.
Country has an impact on Nicotine use.

```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90))
grid.arrange(p1,p2,p3,p4)
```
Ethnicity impacts Nicotine use
NScore has some affect on Nicotine use with most yes observations around 0 and now a little less than 0
EScore appears to have limited to no effect
OScore is similar to NScore with some differences in outlier.


```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```
Ascore has limited effect, similar to OScore and NScore
CScore has limited effect with No median being higher than Yes.
Impulsive and SS have larger quartile ranges compared to other numerical data with SS having more effect on Nicotine.

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()

grid.arrange(p1,p2,p3,p4)
```

## Task 4

Set up our folds for cross-validation  
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```

```{r}
  drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #add tuning of mtry and min_n parameters
  #setting trees to 100 here should also speed things up a bit, but more trees might be better
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)


rf_grid = grid_regular(
  mtry(range = c(2, 8)), #these values determined through significant trial and error
  min_n(range = c(5, 20)), #these values determined through significant trial and error
  levels = 10
)


set.seed(123)
rf_res = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = 20 #try 20 different combinations of the random forest tuning parameters
)
  
```
### Task 5
The most important variables are SS, which is the numeric value for sensation seeing measured by ImpSS or impulsive sensation seeking followed by Country factor level UK meaning living in the United Kingdom is importance to nicotine us.  Finally OScore which measures openness to experience.

```{r}
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```
```{r}
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```
```{r}
best_rf = select_best(rf_res, "accuracy")
view(best_rf)
```

```{r}
final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
```{r}
#fit the finalized workflow to our training data
final_rf_fit = fit(final_rf, train)
```

Check out variable importance
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

### Task6

The model's accuracy against the training data set is higher than on the test set. There's a bigger delta between the accuracy and NRI in the training results vs test results. More concerning is that the accuracy dropped in reviewing new data, which is problematic that it might be better fited on the training set.  Your model assuming all have recent Nicotine use is similar in accuracy to your model against the test data set. 

Predictions  
```{r}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)
```

Confusion matrix
```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```

Predictions on test
```{r}
testpredrf = predict(final_rf_fit, test)
#head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```
### Task 7
Not a very strong mode on the test dataset.  The model seems to be a better fit for the training, which is counter productive using it against new, unseen data.  There's likely limited harm in an inaccurate model predicting nicotine uses based upon psychological and biographical metrics, but there's not much improvement over just assuming everyone uses Nicotine, which is very untrue.
